Human Activity Recognition
==========================

Background
----------

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now 
possible to collect a large amount of data about personal activity 
relatively inexpensively. These type of devices are part of the 
quantified self movement – a group of enthusiasts who take measurements 
about themselves regularly to improve their health, to find patterns 
in their behavior, or because they are tech geeks. One thing that people 
regularly do is quantify how much of a particular activity they do, 
but they rarely quantify how well they do it.

this project we will use data from accelerometers on the belt, forearm, 
arm, and dumbell of 6 participants. They were asked to perform barbell 
lifts correctly and incorrectly in 5 different ways. More information 
is available from the website here: [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har)
(see the section on the Weight Lifting Exercise Dataset). 

Loading and Preprocessing data
------------------------------

#### Loading the data.

The training data for this project are available here: 

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data are available here: 

[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

```{r}
library(ggplot2)
library(caret)
library(randomForest)
library(reshape2)
```

```{r load-data,cache=TRUE}
training <- read.csv('pml-training.csv', na.strings=c('NA', '', '#DIV/0!'))
testing <- read.csv('pml-testing.csv', na.strings=c('NA', '', '#DIV/0!'))
```

#### Preprocessing

remove info features (1 to 7 columns) and features which has NA value.

```{r}
infoList <- 1:7
naList <- which(apply(training, 2, function(x) sum(is.na(x))) > 0)
excludeList <- c(infoList, naList)
training <- training[, -excludeList]
testing <- testing[, -excludeList]
```

split the training data to subTrain and subTest for cross-validation.

```{r}
set.seed(9527)
inTrain <- createDataPartition(training$classe, p=.7, list=FALSE)
subTrain <- training[inTrain,]
subTest <- training[-inTrain,]
```

look up the features correlation.

```{r}
correl <- cor(subTrain[,-dim(subTrain)[2]])
diag(correl) <- 0
g <- ggplot(melt(correl), aes(x=Var1, y=Var2, fill=value))
g <- g + labs(title='Correlation')
g <- g + geom_raster()
g <- g + scale_fill_gradientn(colours=c('red', 'white', 'blue'))
g
```

Fitting and Validation
----------------------

fit model with randomForest

```{r fit-model,cache=TRUE}
set.seed(9527)
fit <- randomForest(classe ~ ., data=subTrain)
fit
```

the OOB estimate of  error rate is 0.61%, it's small enough.

cross-validation

```{r}
confuMatrix <- confusionMatrix(predict(fit, subTest), subTest$classe)
confuMatrix
```

the estimate out of sample error is `r round(1 - confuMatrix$overall[1], 4)*100`% ,
all of class's sensitivity is greater than 99% .

```{r}
cm <- apply(confuMatrix$table, 1, function(x) x/sum(x))
names(dimnames(cm))[1] <- 'Actual'
cm <- melt(cm)
g <- ggplot(cm, aes(x=Prediction, y=Actual, fill=value))
g <- g + labs(title='Confusion Matrix')
g <- g + geom_raster()
g <- g + scale_fill_gradientn(colours=c('white', 'yellow', 'brown', 'black'))
g
```

Conclusions
-----------

We can recognise a small number of exercise mistakes by the measurement from
the accelerometers on the belt, forearm, arm, and dumbell.


Prediction
----------

apply the machine learning algorithm to the testing data set.

```{r}
pred <- predict(fit, testing)
pred
```

```{r eval=FALSE}
pml_write_files = function(x){
	  n = length(x)
  for(i in 1:n){
	      filename = paste0("problem_id_",i,".txt")
      write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
	    }
}
pml_write_files(pred)
```
